---
tittle: rOceansNew - Examples
author: I. Montero-Serra
date: 30/Jul/2018
output: html_document
---

#  Global Biodiversity Patterns in Marine Benthic Ecosystems using rOceans
rOceans is an R Package for exploring spatial trends in marine biodiversity and conservation. It contains functions to access and explore spatial patterns of marine biodiversity. 

## rOceans functions list

(1) *oceanDataAccess()* Access occurrences data from OBIS and GBIF for a list of species and provides options for data quality checks

(2) *oceanDataCheck()* Automatically merges, checks and filters big datasets from OBIS and GBIF

(3) *oceanAbundance()* Computes global or regional spatial grids of abundance of occurrences per grid cell at multiple scales

(4) *oceanMaps()* rapid map visualization tool for representing spatial trends in biodiversity

(5) *oceanDiversity()* computes spatial explicit layers of several diversity metrics including richness, corrected richness, Simpson and Shannon diversity

(6) *oceanEnvironment()* Automatic access to data on global spatial layers of multiple environmental drivers and explores relationships to marine biodiversity metrics

(7) *oceanFuture()* Access current and future environmental data layers and computes expected trends

(8) *oceanHotspots()* classify sites in hotspots and coldspots of biodiversity

(9) *oceanVulnerability()* explores vulnerability of sites by linking biodiversity classifications and expected warming trends


Download rOceansNew from GitHub and load the package
```{r}
invisible(devtools::install_github("monteroserra/rOceans"))
invisible(library(rOceans))

```


## Function #2 *oceanDataCheck* 
Allows merging, checking, filtering the raw data from OBIS and/or GBIF into a common standard database. It removes occurrences with inconsistent or inconmplete taxonomic information, land-based occurrences, and duplicate occurrences. 

For this set of functions, we first  need to access and download data from GBIF (www.gbif.org) and OBIS (www.iobis.org). 

Here I show an example with data of species within the genus *Acropora* that were directly downloaded from the website on June 2018. The data can be accessed from my GitHub repository "monteroserra/rOceanData" or using data()

```{r}
#Accesing raw datasets
data(Acropora_GBIF) 
data(Acropora_OBIS)

#Checking OBIS or GBIF individually
Acropora_OBIS_checked = oceanDataCheck(OBIS_occurrences = Acropora_OBIS, source = "OBIS")

Acropora_GBIF_checked = oceanDataCheck(GBIF_occurrences = Acropora_GBIF, source = "GBIF")
```

Merging OBIS and GBIF Datasets, and checking for taxonomic and geographic issues
```{r}
Acropora_Total_Checked = oceanDataCheck(GBIF_occurrences = Acropora_GBIF,
                                        OBIS_occurrences = Acropora_OBIS,
                                        source = "GBIF_&_OBIS")
```

# Functions for analyzing and visualizing spatial trends in marine biodiversity

## Function #3 *oceanAbundance* 
Allows creating global or delimited spatial grids with abundance of occurrences per cell at differenc cell sizes. 

Following with the Acropora example, we will use this function to compute global rasters of abundance per grid cell for Acropora species at different resolutions (10 x 10) (5 x 5) (1 x 1) and (0.5 x 0.5) using *cell_size*

```{r}
Acropora_abundance_10x10 = oceanAbundance(occurrences = Acropora_Total_Checked, cell_size=10)
Acropora_abundance_5x5 = oceanAbundance(occurrences = Acropora_Total_Checked, cell_size=5)
Acropora_abundance_1x1 = oceanAbundance(occurrences = Acropora_Total_Checked, cell_size=1)
Acropora_abundance_0.5x0.5 = oceanAbundance(occurrences = Acropora_Total_Checked, cell_size=0.5)
```

## Function #3 *oceanMaps*  
A versatile data mapping tool for a rapid visualization of marine biodiversity patterns previously computed.  

```{r}
par(mfrow=c(2,2))
oceanMaps(Acropora_abundance_10x10, logScale=T, main="Acropora abundance (10x10)")
oceanMaps(Acropora_abundance_5x5, logScale=T, main="Acropora abundance (5x5)")
oceanMaps(Acropora_abundance_1x1, logScale=T, main="Acropora abundance (1x1)")
oceanMaps(Acropora_abundance_0.5x0.5, logScale=T, main="Acropora abundance (0.5x0.5)")
par(mfrow=c(1,1))

```
