---
tittle: Exploring Carcharhinidae sharks abundance using from GBIF and OBIS data
author: I. Montero-Serra
date: 30/Jul/2018
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">"
)
```

# Introduction

This document describes how to use rOceans for computing abundance spatial patterns of abundance in marine species using raw data directly downloaded from OBIS and GBIF. Here we have an example of usage for sharks of family Carcharhinidae.  


(1) Download rOceans from GitHub and load the package

```{r,cache.comments=FALSE}
invisible(devtools::install_github("monteroserra/rOceans", force = T))
invisible(library(rOceans))

```


# For simplicity, we provide the dataset "Carcharhinidae_total_checked" which is the output obtained using *oceanDataCheck()* applied to raw datasets from OBIS and GBIF (see vignette 1 with Acropora example for more details on checking and filtering raw data)

```{r}
data(Carcharhinidae_total_checked)

Carcharhinidae_abundance_5 = oceanAbundance(occurrences = Carcharhinidae_total_checked, cell_size=5)

#Visualize abundance
oceanMaps(Carcharhinidae_abundance_5, logScale=T, main="Carcharhinidae abundance")

```

# Computing species abundance matrix - and several diversity metrics
```{r}
Carcharhinidae_diversity = oceanDiversity(occurrences = Carcharhinidae_total_checked, print=F)

# Visualizing diversity metrics
par(mfrow=c(2,2))
oceanMaps(Carcharhinidae_diversity[[2]], main = "Carcharhinidae richness")
oceanMaps(Carcharhinidae_diversity[[3]], main = "Carcharhinidae Corrected richness Diversity")
oceanMaps(Carcharhinidae_diversity[[4]], main = "Carcharhinidae Shannon Diversity")
oceanMaps(Carcharhinidae_diversity[[5]], main = "Carcharhinidae Simpson Diversity")

par(mfrow=c(1,1))
```
